---
title: "Reproducible Research: Peer Assessment 1-Victor Vaquero Soto"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, destfile ="./activity.zip",method="curl")
unzip(zipfile = "./activity.zip", exdir = "./")
data<-read.csv("activity.csv",sep=",")
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$monthday<-(format(data$date, "%m-%d"))
str(data)
```


## What is mean total number of steps taken per day?
```{r perday, echo=TRUE,fig.height=3}
library(ggplot2)
total<-with(data,tapply(steps, monthday, sum,na.rm=TRUE))
par(mar=c(5,5,3,3))
qplot(unique(data$monthday),as.vector(total), show.legend = FALSE)+geom_bar(stat = "identity", show.legend = FALSE)+scale_x_discrete(breaks=c("10-15","10-31","11-15")) +
  xlab("Day") + ylab("Total steps") +
  ggtitle("Total number of steps per day") + 
  theme(plot.title = element_text(hjust = 0.5))
  total<-with(data,tapply(steps, monthday, sum))
  meansteps<-mean(total,na.rm=TRUE)
  mediansteps<-median(total,na.rm=TRUE)
```
The mean is `r meansteps` and the median is `r mediansteps`.

## What is the average daily activity pattern?
```{r perinterval, echo=TRUE,fig.height=5}
averageperinterval<-aggregate(steps ~ interval, data = data, FUN = mean, na.rm = TRUE)
par(mar=c(5,5,3,3))
plot(unique(data$interval),averageperinterval[,2],type="l",col="red",
       main = "Average daily activity pattern",  
       xlab = "5-minute Intervals",  
       ylab = "Average Steps Taken")
maximum<-averageperinterval[which.max(averageperinterval$steps),1]
``` 
The maximum of 5-minute interval is `r maximum`.


## Imputing missing values
```{r perdaynoNAs, echo=TRUE,fig.height=3}
nofNA<-colSums(is.na(data))
print(nofNA)
newdata <- data
nas<-is.na(newdata$steps)
meanperinterval<-with(newdata,tapply(steps, interval, mean,na.rm=TRUE))

newdata$steps[nas]<-meanperinterval[as.character(newdata$interval[nas])]

total<-with(newdata,tapply(steps, monthday, sum,na.rm=TRUE))
par(mar=c(5,5,3,3))
qplot(unique(newdata$monthday),as.vector(total), show.legend = FALSE)+geom_bar(stat = "identity", show.legend = FALSE)+scale_x_discrete(breaks=c("10-15","10-31","11-15")) +
  xlab("Day") + ylab("Total steps") +
  ggtitle("Total number of steps per day without NA") + 
  theme(plot.title = element_text(hjust = 0.5))

  newmeansteps<-mean(total,na.rm=TRUE)
  newmediansteps<-median(total,na.rm=TRUE)
```
The number of NA in steps is  `r nofNA[1]`.

The new mean is `r newmeansteps` and the new median is `r newmediansteps`. As we can see the mean is the same to the one of the data set with NA while the medians are a bit different. Of course the quantiles are different.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekend, echo=TRUE,fig.height=5}
library(lattice)
Sys.setlocale(locale = 'en_US.UTF-8')
newdata$day <- ifelse(weekdays(newdata$date) == "Saturday" | weekdays(newdata$date) == "Sunday", "weekend", "weekday")
averageperintervalandday<-aggregate(steps ~ interval+day, data = newdata, FUN = mean, na.rm = TRUE)
xyplot(steps~interval|day, data=averageperintervalandday, type="l",  layout = c(1,2),  ylab="Average Number of Steps per Interval", xlab="Interval")
```

